---
title: "R Notebook"
output: html_notebook
---

```{r}
install.packages("ca")
library(ca)
```

Data for the lecture

```{r}
data_vec <- c(1933, 1175, 1186, 646, 579, 671, 707, 780, 767, 768, 962, 1126)
data_mat <- matrix(data = data_vec, nrow = 4, ncol = 3, byrow = T)
rownames(data_mat) <- c("Rural", "city 20k", "city 20k-100k", "city 100k+")
colnames(data_mat) <- c("Realised","Refusals","Errors")
data_mat
```

In order to verify whether the correspondence analysis may be applied one should follow the steps:

1. verify relationship (using $\chi^2$ test)
2. verify correlation (using Cramers' V statistic)
3. proceed with correspondence analysis

```{r}
chisq.test(data_mat)
```

To conduct the correspondence analysis in R we can use function `ca` from the package `ca`.

```{r}
result <- ca(data_mat)
plot(result)
```

```{r}
result
```

Follow the example about ages and tv stations

```{r}
vec <- c(26, 10, 40, 44, 18, 14, 15, 12, 21)
tab <- matrix(data = vec, nrow = 3, byrow = T)
colnames(tab) <- c("tvp1", "polsat", "tvn")
rownames(tab) <- c("<20","20-35", "35+")
chisq.test(tab)
```
```{r}
sum(tab)
```

```{r}
result_tv <- ca(tab)
result_tv
```

```{r}
plot(result_tv)
```

