---
author: "Maciej Beręsewicz"
title: "Correspondence Analysis"
output: html_notebook
---

# Setting for notebook

```{r}
library(reticulate)
use_python("/usr/local/anaconda3/bin/python")
```

```{r}
knitr::opts_chunk$set(engine.path = list(
  python = "/usr/local/anaconda3/bin/python",
  julia = "/Applications/Julia-1.3.app/Contents/Resources/julia/bin/"
))
```

# Correspondence analysis

## Correspondence analysis in R

```{r}
install.packages("ca")
library(ca)
```

Data for the lecture

```{r}
data_vec <- c(1933, 1175, 1186, 646, 579, 671, 707, 780, 767, 768, 962, 1126)
data_mat <- matrix(data = data_vec, nrow = 4, ncol = 3, byrow = T)
rownames(data_mat) <- c("Rural", "city 20k", "city 20k-100k", "city 100k+")
colnames(data_mat) <- c("Realised","Refusals","Errors")
data_mat
```

In order to verify whether the correspondence analysis may be applied one should follow the steps:

1. verify relationship (using $\chi^2$ test)
2. verify correlation (using Cramers' V statistic)
3. proceed with correspondence analysis

```{r}
chisq.test(data_mat)
```

To conduct the correspondence analysis in R we can use function `ca` from the package `ca`.

```{r}
result <- ca(data_mat)
plot(result)
```

```{r}
result
```

Follow the example about ages and tv stations

```{r}
vec <- c(26, 10, 40, 44, 18, 14, 15, 12, 21)
tab <- matrix(data = vec, nrow = 3, byrow = T)
colnames(tab) <- c("tvp1", "polsat", "tvn")
rownames(tab) <- c("<20","20-35", "35+")
chisq.test(tab)
```
```{r}
sum(tab)
```

```{r}
result_tv <- ca(tab)
result_tv
```

```{r}
plot(result_tv)
```

## Correspondence analysis in Python

To conduct correspondence analysis we need to install module `prince` (https://github.com/MaxHalford/prince) or we may implement the correspondence analysis by ourselves

```{python}
import prince
import pandas as pd
```

Take the data

```{python}
X = pd.DataFrame(
    data=[
        [1933, 1175, 1186],
        [646, 579, 671],
        [707, 780, 767],
        [768, 962, 1126]
    ],
    columns=pd.Series(["Realised","Refusals","Errors"]),
    index=pd.Series(["Rural", "city 20k", "city 20k-100k", "city 100k+"])
)
X
```

Define settings for correspondence analysis

```{python}
ca = prince.CA(n_components=2, n_iter=100, copy=True,check_input=True,engine='auto',random_state=42)
ca = ca.fit(X)
ca
```

```{python}
dir(ca)
```

The main important elements

col_masses_
row_masses_
column_coordinates
row_coordinates
eigenvalues_
explained_inertia_
total_inertia_

```{python}
print("---------- row coords ----------")
ca.row_coordinates(X)
print("---------- col coords ----------")
ca.column_coordinates(X)
print("---------- col masses ----------")
ca.col_masses_
print("---------- row masses ----------")
ca.row_masses_
print("---------- % of intertia explained ----------")
ca.explained_inertia_
```


```{python}
ca.plot_coordinates(X = X).get_figure()
```


## Correspondence analysis in Julia

We need to implement correspondence analysis by ourselves

