---
title: "R Notebook"
output: html_notebook
---

```{r}
install.packages("margins")
library(margins)
```

```{r}
mtcars
```

Let's assume that we would like to build a model where 

            mpg ~ gear 

and we assume that gear is categorical

$$
mpg = \beta_0 + \beta_1 \times gear
$$

```{r}
m1 <- lm(formula = mpg ~ gear, data = mtcars)
summary(m1)
```

```{r}
table(mtcars$gear)
```

To inform R that gear is categorical variable we need to use function factor 

```{r}
mtcars$gear_f <- factor(mtcars$gear)
mtcars
```

Now, we will include gear_f into the model

$$
mpg = \beta_0 + \beta_1 \times (gear = 4) + \beta_2 \times (gear = 5)
$$

```{r}
m2 <- lm(formula = mpg ~ gear_f, data = mtcars)
summary(m2)
```

```{r}
aggregate(mpg ~ gear, data = mtcars, FUN = mean)
```

How to change the contrast

```{r}
m3 <- lm(mpg ~ gear_f, data  = mtcars, contrast = list(gear_f = "contr.SAS"))
summary(m3)
```

In order to change the reference level we can use function relevel (reference level).

```{r}
m4 <- lm(formula = mpg ~ relevel(x = gear_f, ref = "4"), data = mtcars)
summary(m4)
```

How to check what is the reference level 

```{r}
levels(mtcars$gear_f)
```

```{r}
mtcars$gear_f2 <- relevel(mtcars$gear_f, ref = "4")
levels(mtcars$gear_f2)
```

Now, we will use contr.sum to compare to overall mean

```{r}
m5 <- lm(formula = mpg ~ gear_f, data = mtcars, contrast = list(gear_f = "contr.sum"))
summary(m5)
```

```{r}
result <- aggregate(mpg ~ gear, data = mtcars, FUN = mean)[, "mpg"]
result
mean(result)
```

Margial effects

x1*x2 = x1 + x2 + x1:x2

```{r}
m6_1 <- lm(mpg ~ wt + gear_f + am, data = mtcars)
m6_2 <- lm(mpg ~ wt + I(wt^2) + gear_f + am, data = mtcars)
m6_3 <- lm(mpg ~ wt + I(wt^2) + gear_f*am, data = mtcars)

summary(m6_2)
```

Calculate **Average marginal effects** for model m6_1

```{r}
margins(m6_1)
summary(margins(m6_2))
```

Average marginal effect for model 

$$
Y = \beta_0 + \beta_1X_1 + \beta_2X_2 + \beta_3 X_1X_2
$$

$$
\frac{\partial Y}{\partial X_1} = \beta_1 + \beta_3 X_2
$$

$$
\frac{\partial Y_i}{\partial X_{i1}} = \beta_1 + \beta_3 X_{i2}
$$

and then we calculate Average Marginal Effects as follows

$$
AME = \frac{\sum_{i} \hat{\beta}_1+ \hat{\beta}_3X_{i2}}{n}
$$

Marginal effects at means (MEMs) as

$$
MEMs = \hat{\beta}_1 + \hat{\beta}_3 \bar{X}_{2}
$$





