---
author: "Maciej Beręsewicz"
title: "R Notebook"
output: html_notebook
---

Load the packages

```{r}
#install.packages("speedglm")
#install.packages(c("mfx", "BaylorEdPsych", "ROCR"))
library(tidyverse)
library(speedglm) ## high-performance GLM models
library(mfx) ## marginal effects for GLM models
library(BaylorEdPsych) ## pseudoR2 measures for GLM
library(ROCR) ## for prediction
```

load the data

```{r}
df <- readRDS("../data/data_for_lecture.rds") 
df$locality <- as_factor(df$locality)
df$woj <- as_factor(df$woj)
df$gender <- as_factor(df$gender)
df

table(as_factor(df$status))

df 
```

We build a model only with gender (glm or speedglm)

```{r}
m1 <- glm(formula = status ~ gender, data = df, family = binomial(link = "logit"))
summary(m1)
```

$$
P( \text{status} = 1 | \text{gender} ) = \frac{\exp(-0.65479 + 0.12165 \cdot \text{Kobieta})}{1 + \exp(-0.65479 + 0.12165 \cdot \text{Kobieta})}
$$

Calculate odds ratio

```{r}
round(exp(coef(m1)),2)
```

We calculate 95% confidence interval 

```{r}
round(exp(confint(m1)),2)
```

Let's add age to our model

```{r}
m2 <- glm(formula = status ~ gender + age, data = df, family = binomial(link = "logit"))
summary(m2)
```
```{r}
round(coef(m2),2)
```

$$
P( \text{status} = 1 | \text{gender} ) = 
\frac{\exp(-1.01 + 0.13 \cdot \text{Kobieta} + 0.01 \cdot \text{age})}
{1 + \exp(-1.01 + 0.13 \cdot \text{Kobieta}  + 0.01 \cdot \text{age})}
$$

We calculate odds ratio to compare 20s with 50s

```{r}
exp(0.006496348*(50-20))
exp(0.006496348*(70-20))
```

```{r}
plot(20:70, exp(0.006496348*(20:70-20)), xlab = "Age", ylab = "Odds Ratios")
```


```{r}
m3 <- speedglm(formula = status ~ gender + age, data = df, family = binomial(link = "logit"))
summary(m3)
```

Compare two models based on AIC, BIC -- the lower the better

$$
\text{AIC} = -2 \cdot LL + 2 \cdot k
$$

$$
\text{BIC} = -2 \cdot LL + k \cdot \log(n)
$$


```{r}
AIC(m1, m2)
BIC(m1, m2)
```



Add to the model locality -- odds ratio, AIC

```{r}
m3 <- update(m2, . ~ . + locality)
AIC(m1, m2, m3)
BIC(m1, m2, m3)
```

```{r}
summary(m3)
```

```{r}
exp(coef(m3))
```

Compare two models

```{r}
anova(m1, m2, m3, test = "LRT")
```

Calculate McFadden's pseudo R2 -- pseudoR2 (pakiet -- `BaylorEdPsych`)

```{r}
1 - 14437/14751
```

Create new model

```{r}
m4 <- glm(formula = status ~ gender + age + locality + woj, 
          data = df, 
          family = binomial(link = "logit")) ## or family = "binomial"
summary(m4)
```

Calculate Pseudo R2 measures using BaylorEdPsych package

```{r}
measures <- PseudoR2(m4)
measures[1:4]
```

How good is the model on predicting our Y variable

```{r}
pred_m4 <- predict(m4, newdata = df) ## by default predict function returns eta
head(pred_m4)
pred_m4 <- predict(m4, newdata = df, type = "response") ## probability
head(pred_m4)
hist(pred_m4, breaks = "scott")
```

In order to predict the class (Y = 1 or Y = 0) we need to apply some threshold, by default we can use 0.5.

Confusion Matrix -- in rows we have observed values of Y and in columns we have predictions

```{r}
conftab1 <- table(df$status == 1, pred_m4 >= 0.5)
conftab1
```

- TP = 485, TN = 6817, FP = 429, FN = 3569
- Accuracy: (TP+TN)/Total

```{r}
(485+6817) / sum(conftab1)*100
```

- Misclassification Rate: (FP+FN)/Total

```{r}
(429+3569)/sum(conftab1)*100
```

- True Positive Rate - TP/ actual yes

```{r}
485/(485+3569)
```

- False Positive Rate - FP/actual no

```{r}
429/(429+6817)
```

- Specificity -- TN/actual no

```{r}
6817/(6817 + 429)
```

- Precision -- TP/predicted yes

```{r}
485/(485 + 429)
```

1. problem is our sample -- it is not balanced (not 50/50)
2. if our sample is unbalanced then using threshold 0.5 is meaningless -- we should use value that is closer to share of Ys==1 in our data.
3. there is a room for optimization -- select threshold to maximize selected measures

```{r}
conftab1 <- table(df$status == 1, pred_m4 >= 0.50)
conftab2 <- table(df$status == 1, pred_m4 >= 0.35)

prop.table(conftab1, margin = 1)
prop.table(conftab2, margin = 1)
```

Now proceed to ROCR package

```{r}
pred <- prediction(predictions = pred_m4, labels = as.numeric(df$status))
pred
perf <- performance(prediction.obj = pred,"tpr","fpr")
perf
plot(perf)
```

## Calculate marginal effects

```{r}
mf_model1 <- mfx::logitmfx(formula = status ~ age + gender + locality, data = df)
mf_model1
```

```{r}
mf_model2 <- mfx::logitor(formula = status ~ age + gender + locality, data = df)
mf_model2
```


```{r}
ods_model2 <- glm(formula = status ~ age + gender + locality, data = df, family = binomial)
exp(coef(ods_model2))
```

-- LASSO -- glmnet -- when Xs are correlated, when # variables > # cases, for prediction
-- logistf -- perfect separation in logistic regression
-- rms package -- regression modelling strategies






