---
author: "Maciej Beręsewicz"
title: "R Notebook"
output: html_notebook
---

Load the packages

```{r}
library(tidyverse)
#install.packages("speedglm")
library(speedglm)
```

load the data

```{r}
df <- readRDS("../data/data_for_lecture.rds") 
df$locality <- as_factor(df$locality)
df$woj <- as_factor(df$woj)
df$gender <- as_factor(df$gender)
df

table(as_factor(df$status))

df 
```

We build a model only with gender (glm or speedglm)

```{r}
m1 <- glm(formula = status ~ gender, data = df, family = binomial(link = "logit"))
summary(m1)
```

$$
P( \text{status} = 1 | \text{gender} ) = \frac{\exp(-0.65479 + 0.12165 \cdot \text{Kobieta})}{1 + \exp(-0.65479 + 0.12165 \cdot \text{Kobieta})}
$$

Calculate odds ratio

```{r}
round(exp(coef(m1)),2)
```

We calculate 95% confidence interval 

```{r}
round(exp(confint(m1)),2)
```

Let's add age to our model

```{r}
m2 <- glm(formula = status ~ gender + age, data = df, family = binomial(link = "logit"))
summary(m2)
```
```{r}
round(coef(m2),2)
```

$$
P( \text{status} = 1 | \text{gender} ) = 
\frac{\exp(-1.01 + 0.13 \cdot \text{Kobieta} + 0.01 \cdot \text{age})}
{1 + \exp(-1.01 + 0.13 \cdot \text{Kobieta}  + 0.01 \cdot \text{age})}
$$

We calculate odds ratio to compare 20s with 50s

```{r}
exp(0.006496348*(50-20))
exp(0.006496348*(70-20))
```

```{r}
plot(20:70, exp(0.006496348*(20:70-20)), xlab = "Age", ylab = "Odds Ratios")
```


```{r}
m3 <- speedglm(formula = status ~ gender + age, data = df, family = binomial(link = "logit"))
summary(m3)
```

Compare two models based on AIC, BIC -- the lower the better

$$
\text{AIC} = -2 \cdot LL + 2 \cdot k
$$

$$
\text{BIC} = -2 \cdot LL + k \cdot \log(n)
$$


```{r}
AIC(m1, m2)
BIC(m1, m2)
```



Add to the model locality -- odds ratio, AIC

```{r}
m3 <- update(m2, . ~ . + locality)
AIC(m1, m2, m3)
BIC(m1, m2, m3)
```

```{r}
summary(m3)
```

```{r}
exp(coef(m3))
```

Compare two models

```{r}
anova(m1, m2, m3, test = "LRT")
```

Calculate McFadden's pseudo R2 -- pseudoR2 (pakiet -- `BaylorEdPsych`)

```{r}
1 - 14437/14751
```


Build a model

```{r}
m1 <- glm(formula = status ~ gender + age + locality + woj, 
          data = df, 
          family = binomial(link = "logit")) ## or family = "binomial"
summary(m1)
```

```{r}
m1_marg <- margins(m1)
plot(m1_marg)
```

```{r}
plot(ggeffect(m1))
```



