---
title: "R Notebook"
output: html_notebook
---


```{r}
install.packages("countreg", repos="http://R-Forge.R-project.org")
install.packages("AER")
library(haven) ## for reading sas files
library(MASS) ## for glm.nb -- negative binomial
library(vcd) ## for testing goodnes of fit for poisson and negative binomial
library(countreg) ## for modelling count data
library(AER) ## bunch of useful functions

df <- read_sas("../data/baza2005.sas7bdat")
head(df)
```

1. start with visualisation

```{r}
barplot(table(df$LOS_LT14), xlab = "Number of children under 14", 
        ylab = "Number of households")
```

2. Is it a Poisson distribution?

```{r}
mean(df$LOS_LT14) ## expected value E(Y) = lambda
var(df$LOS_LT14) ## variance of Y Var(Y) > E(Y) -> overdispersion

var(df$LOS_LT14)/mean(df$LOS_LT14)
```


3. We may do the tests before modeling 

```{r}
plot(goodfit(df$LOS_LT14, type = "poisson"))
plot(goodfit(df$LOS_LT14, type = "nbinomial", par = list(size = 10)))
```

4. Now, we will create three model:

+ poisson  (use glm function)
+ quasi-poisson  (use glm function)
+ negative binomial (using MASS::glm.nb function)


$$
\text{LOS_LT14}  = \exp(\beta_1 + \beta_2 \text{DOCH} + ...)
$$

$$
\log(\text{LOS_LT14}) = \beta_1 + \beta_2 \text{DOCH} + ...
$$


```{r}
m1 <- glm(formula = LOS_LT14 ~ I(DOCH / 1000) + factor(KLM) + factor(WOJ),
          data = df,
          family = poisson())
summary(m1)
```

Używamy transformacji odwrotnej do logarytmu naturalnego czyli funkcję e

```{r}
exp(coef(m1))
```

```{r}
m2 <- glm(LOS_LT14 ~ I(DOCH/1000) + factor(KLM) + factor(WOJ), 
         data = df,
         family = quasipoisson())
summary(m2)
```

To interpret the parameters from the glm model we just need to use the exp transformation (because by default in Poisson and quasi-Poisson we use log link)

```{r}
exp(coef(m1)) 
```

We use dispersion test to verify if dispersion is actually the problem


$$
\begin{cases}
trafo(\mu) = \mu, & \text{if quasi-poisson,} \\ 
trafo(\mu) = \mu^2, & \text{if negative binomial,}
\end{cases}
$$

$$
\begin{cases}
Var[y] = \mu + \alpha \cdot \mu, & \text{if quasi-poisson,} \\ 
Var[y] = \mu + \alpha \cdot \mu^2, & \text{if negative binomial,}
\end{cases}
$$

```{r}
dispersiontest(m1, trafo = 1) ## quasipoisson / negative binomial type 1
dispersiontest(m1, trafo = 2) ## negative binomial type 2
```

Negative-binomial model

```{r}
m3 <- glm.nb(formula = LOS_LT14 ~ I(DOCH/1000) + factor(KLM) + factor(WOJ), 
             data = df)
summary(m3)
```

Jak porównać te dwa modele -- poisson i negative binomial

```{r}
data.frame(poisson = exp(coef(m1)), negbin = exp(coef(m3)))
```

Compare which model is better -- Poisson or Negative-Binomial

```{r}
AIC(m1,m3)
BIC(m1,m3)
```

```{r}
countreg::rootogram(m1, main = "Poisson")
countreg::rootogram(m3, main = "Negative-Binomial")
```

1. visualize
2. calculate the models (poisson, quasipo, negbin, ...)
3. assess the models
4. select the "right" model (according to some measures)

