---
title: "R Notebook"
output: html_notebook
---

Load packages

```{r}
#install.packages(c("geepack", "MuMIn"))
library(geepack) ## to estimate GEE models
library(MuMIn) ## selection of GEE models -- we need to choose which R matrix to apply
library(readxl) ## read excel files library 
library(lubridate)
library(tidyverse)
```

Read the data from excel file and sort data by id (customer) i data (date of transaction)

```{r}
piwa <- read_excel("../data/piwa.xlsx") %>%
  mutate(weekd = wday(ymd(data),week_start = 1)) %>%
  arrange(id, data) %>%
  add_count(id, name = "times")  %>%
  filter(times > 2, times <= 10)

head(piwa)

piwa %>%
  group_by(id) %>%
  summarise(numer_of_pur = n(),
            beers_sum = sum(sztuki),
            beers_mean = mean(sztuki),
            beers_sd = sd(sztuki))
```

```{r}
piwa %>%
  count(data = ymd(data), wt = sztuki) %>%
  ggplot(data = ., aes(x = data, y = n)) +
  #geom_point() +
  geom_line()
```

Wizualizacja liczby zakupionych piw

```{r}
barplot(table(piwa$sztuki))
```

```{r}
m0 <- glm(formula = sztuki ~ wielkosc_gosp + factor(weekd) + klm + factor(miesiac) + 
            sklepy + wyksztalcenie + wojewodztwo,  
          data = piwa,
          family  = quasipoisson())

summary(m0)
```


```{r}
m1 <- geeglm(formula = sztuki ~ wielkosc_gosp + factor(weekd)+ klm + factor(miesiac) + 
               sklepy + wyksztalcenie  + wojewodztwo, 
             data = piwa,
             family  = poisson(), 
             corstr = "independence",
             id = id, ## we need to sort data according to (id and) time!!!!
             scale.fix = FALSE)

summary(m1)
```

Oszacujemy model, który zakłada, że nasze dane są skorelowanie i określimy R

Parametry modelu:

corstr -- określenie struktury macierzy R  "independence"', '"exchangeable"', '"ar1"', '"unstructured"', '"userdefined"'
id  -- identyfikator jednostki
scale.fix -- parametr phi = 1
 

```{r}
m3 <- geeglm(formula = sztuki ~ wielkosc_gosp + factor(weekd) + klm + 
               factor(miesiac) + sklepy + wyksztalcenie + wojewodztwo, 
             data = piwa,
             family  = poisson(), ## quasipoisson
             corstr = "exchangeable", ## only one alpha
             id = id,
             scale.fix = FALSE)

m5 <- geeglm(formula = sztuki ~ wielkosc_gosp + factor(weekd) + klm + 
               factor(miesiac) + sklepy + wyksztalcenie + wojewodztwo, 
             data = piwa,
             family  = poisson(),
             corstr = "ar1", ## autocorrelation level 1
             id = id,
             scale.fix = FALSE)

summary(m3)
```


Porównamy model m0 z modelem m2

```{r}
model.sel(m1, m3, m5, rank = "QIC")
```

```{r}
data.frame(m1 = coef(m1), m3 = coef(m3))
```



